---
title: "Lab 8: Data Frame Functions and Simulation Exploration"
author: "Aiden Rodriguez"
format: 
  html:
    embed-resources: true
    code-tools: true
    toc: true
    html-table-processing: none
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---


```{r}
library(tidyverse)
library(knitr)
```

```{r}
fish = read_csv("BlackfootFish.csv")
```
### Standardizing Variables

```{r}
#| echo: true
#| eval: false
fish <- fish |> 
  mutate(length = (length - min(length, na.rm = TRUE)) / 
           (max(length, na.rm = TRUE) - min(length, na.rm = TRUE)), 
         weight = (weight - min(weight, na.rm = TRUE)) / 
           (max(weight, na.rm = TRUE) - min(length, na.rm = TRUE)))
```

**1. What is the mistake I made in the above rescaling code?**

The first one does max length - min length, however the second weight based one does max weight - min length. It should both be weight here and was a copy paste error that was pretty easy to overlook.

**2. Start by writing a function that rescales a vector. Transform the repeated process above into a `rescale_01()` function. Your function should...**

```{r}

rescale_01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

```


**3. Run the code below to test your function. Verify that the maximum of your rescaled vector is 1 and the minimum is 0!**

```{r}
rescaled <- rescale_01(fish$weight)

min(rescaled, na.rm = TRUE)
max(rescaled, na.rm = TRUE)
```


**4. Create a `rescale_column()` function that accepts two arguments:**

```{r}

rescale_column <- function(df, cols) {
  df |> 
    mutate(
      across({{ cols }}, rescale_01)
    )
}

```

**5. Run the code below to test your function. You might need to edit the code if you call your dataset something different.**

```{r}
fish |> 
  rescale_column(c(length, weight)) |> 
  select(length:weight) |> 
  slice_head(n = 5) |> 
  kable()
```

### Accounting for Unlikely Values

**6. Write a function that will replace unlikely values of a variable in a dataframe with missing values. Your function should accept at least four inputs**:

```{r}
clean_values <- function(df, col, min_val, max_val) {
  df |> 
    mutate(
      {{ col }} := if_else(
        {{ col }} < min_val | {{ col }} > max_val,
        NA,
        {{ col }}
      )
    )
}
```

**7. Write a test that demonstrates that your function worked using the fish data!**

```{r}
clean_values(fish, length, 100, 300)
# Here I just use 300 as a value to show that it will remove some max values. 
# But it seems like the actual reasonable limit is something like 400.
```

## Random Babies Simulation

**8. Simulate the distribution of the number of babies that are correctly returned if there were four babies born in a night at our disorganized hospital. Use 10,000 simulations. Make sure to add a line of code to make your simulation reproducible every time you run it.**


```{r}
#| label: function-for-random-babies

randomBabies <- function(babies){
  
  correct_order <- 1:babies
  returned_order <- sample(1:babies)
  sum(correct_order == returned_order)
  
}
```

```{r}
#| label: full-simulation-for-random-babies

results <- map_int(
  .x = 1:10000,
  .f = ~ randomBabies(4)
)

```

**9. Create a table displaying the proportion of simulations where 0, 1, 2, 3, and 4 babies were given to their correct parent(s).** 


```{r}
#| label: table-for-random-babies

raw_df <- enframe(results, value = "matches")

proportions <- raw_df |>
  count(matches) |>
  mutate(proportion = n / sum(n))

kable(proportions)

```

**10. Now create a barplot showing the proportion of simulations where 0, 1, 2, 3, and 4 babies were given to their correct parent(s).** Don't forget a title and appropriate axis labels.


```{r}
#| label: visualization-for-random-babies

ggplot(proportions, aes(x = matches,
                        y = proportion,
                        fill = factor(matches))) +
  geom_col() +
  labs(
    title = "Distribution of Correctly Returned Babies",
    x = "Number of Correctly Returned Babies",
    y = "Proportion of Correctly Returned Babies",
    fill = "Correct Matches"
  )
```

