---
title: "Stat 331/531: Final Short Answer"
date: now
author: "Aiden Rodriguez"
format: 
  pdf:
    embed-resources: true
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r setup}
library(tidyverse)
library(knitr)
library(maps)
library(mapproj)
```


trail_id: An ID number for the trail
name: Name of the trail
area_name: National Park where the trial is located
city_name: City where the trail is located
state_name: State where the trail is located
country_name: Country where the trail is located
_geoloc: Latitude and longitude of trail
popularity: A score from 0 to 100 of how often the trail is used.
length: Distance covered by trail (in meters).
elevation_gain: How much uphill is on the trail (in meters).
difficulty_rating: Average user rating of trail difficulty
route_type: Type of trail (loop, out and back, or point to point)
visitor_usage: (Honestly, I’m not sure what this one is.)
avg_rating: The average visitor rating of the trail (1 to 5 stars)
num_reviews: The total number of reviews on the app for that trail
features: Notable elements about the trail
activities: Common uses for the trail
units: This variable is a lie! All the lengths are in meters.

```{r}
park_data  <- read_csv("national_park_trails.csv")
```

**Question 1.**

```{r}
park_data <- read_csv("national_park_trails.csv")

park_data_clean <- park_data |>
  # Parse _geoloc JSON string into lat/lng columns
  mutate(
    latitude  = as.numeric(str_extract(`_geoloc`, "(?<='lat': )[0-9.-]+")),
    longitude = as.numeric(str_extract(`_geoloc`, "(?<='lng': )[0-9.-]+"))
  ) |>
  select(-`_geoloc`) |>
  # Normalize difficulty_rating to 0–1
  mutate(
    difficulty_rating = (difficulty_rating - min(difficulty_rating, na.rm = TRUE)) /
                        (max(difficulty_rating, na.rm = TRUE) - min(difficulty_rating, na.rm = TRUE))
  )
```

```{r}
park_data_clean |>
  mutate(
    features = str_extract_all(features, "(?<=')[^']+(?=')")
  ) |>
  unnest(features) |>
  filter(features == "waterfall")
```

Using map() from the purrr package, split park_data_clean by state_name and fit a linear model predicting avg_rating from difficulty_rating and length for each state. Extract the R² value from each model and return a tidy dataframe showing each state alongside its R², sorted from best to worst fit.
```{r}
thresholds <- seq(0, 20, by = 2)

correlations <- map_dbl(.x = thresholds,
                        .f = ~ cor(
                          filter(park_data_clean, popularity > .x)$difficulty_rating,
                          filter(park_data_clean, popularity > .x)$avg_rating,
                          use = "complete.obs"
                        ))

results <- tibble(threshold = thresholds,
                  correlation = correlations)

results

ggplot(results, aes(x = threshold, y = correlation)) +
  geom_line(color = "#378ADD", linewidth = 0.8) +
  geom_point(color = "#378ADD", size = 2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "darkgrey") +
  labs(
    x     = "Minimum popularity threshold",
    y     = "Correlation (difficulty vs. avg rating)",
    title = "Correlation between difficulty and rating\nat varying popularity thresholds"
  ) +
  theme_minimal()
```
Simulation question:
A hiker randomly selects a trail each day from park_data_clean. They will keep hiking as long as every trail they've hiked so far has had a difficulty_rating (normalized 0–1) below 0.75. The moment they hit a trail at or above 0.75, they stop for the season.
Using a simulation with 10,000 iterations:

Simulate how many trails the hiker completes before stopping (or exhausting all trails)
Plot the distribution of trails completed per season as a histogram
Report the mean and 95th percentile of trails completed

Hint: use sample() to draw trails with replacement each iteration, and which() or a while loop to find when the first hard trail appears.
```{r}
set.seed(123)
n_sims <- 10000

# Pull the difficulty ratings as a vector
difficulties <- park_data_clean$difficulty_rating

simulate_season <- function() {
  # Sample trail difficulties with replacement
  sampled <- sample(difficulties, size = length(difficulties), replace = TRUE)
  
  # Find index of first trail at or above 0.75
  hard_trail <- which(sampled >= 0.75)[1]
  
  # If no hard trail found, hiker completes all; otherwise trails completed = index - 1
  if (is.na(hard_trail)) {
    return(length(sampled))
  } else {
    return(hard_trail - 1)
  }
}

results <- replicate(n_sims, simulate_season())

# Summary stats
cat("Mean trails completed:      ", round(mean(results), 2), "\n")
cat("95th percentile completed:  ", quantile(results, 0.95), "\n")

# Histogram
tibble(trails_completed = results) |>
  ggplot(aes(x = trails_completed)) +
  geom_histogram(binwidth = 1, fill = "#378ADD", color = "white", linewidth = 0.2) +
  geom_vline(xintercept = mean(results),
             color = "#A32D2D", linewidth = 0.8, linetype = "dashed") +
  annotate("text",
           x     = mean(results) + 1,
           y     = Inf,
           label = paste("mean =", round(mean(results), 1)),
           vjust = 2, hjust = 0,
           color = "#A32D2D", size = 3.5) +
  labs(
    x     = "Trails completed before stopping",
    y     = "Count",
    title = "Distribution of trails hiked per season"
  ) +
  theme_minimal()
```

```{r}
state_ratings <- park_data_clean |>
  mutate(region = str_to_lower(state_name)) |>
  group_by(region) |>
  summarise(avg_rating = mean(avg_rating, na.rm = TRUE), .groups = "drop")

state_map <- map_data("state")

state_difficulty <- state_map |>
  left_join(state_ratings, by = "region")

# State borders layer
state_borders <- state_map |>
  group_by(region) |>
  summarise(lat = mean(lat), long = mean(long))

ggplot(state_difficulty,
       aes(x = long, y = lat, group = group, fill = avg_rating)) +
  geom_polygon(color = "white", linewidth = 0.1) +
  geom_polygon(data = state_map,
               aes(x = long, y = lat, group = group),
               fill = NA, color = "#222222", linewidth = 0.4) +
  scale_fill_gradient(
    low      = "#C0DD97",
    high     = "#0C447C",
    na.value = "grey90",
    name     = "Avg. Trail Rating"
  ) +
  labs(x = "", y = "") +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8)
  )
```

```{r}
county_map <- map_data("county")

# Prepare trail data for joining - normalize names to match map_data format
trail_county_data <- park_data_clean |>
  mutate(
    region    = str_to_lower(state_name),
    subregion = str_to_lower(city_name)
  ) |>
  group_by(region, subregion) |>
  summarise(difficulty_rating = mean(difficulty_rating, na.rm = TRUE), .groups = "drop")

county_difficulty <- county_map |>
  left_join(trail_county_data, by = c("region", "subregion"))

# Plot
ggplot(county_difficulty,
       aes(x = long, y = lat, group = group, fill = difficulty_rating)) +
  geom_polygon(color = "darkgrey", linewidth = 0.1) +
  scale_fill_gradient(
    low      = "#C0DD97",
    high     = "#A32D2D",
    na.value = "white",
    name     = "Difficulty\n(normalized)"
  ) +
  labs(x = "", y = "") +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  theme_void() +
  theme(
    legend.position  = "right",
    legend.title     = element_text(size = 9),
    legend.text      = element_text(size = 8)
  )
```

```{r}
#This is the important shid

cv_it <- function(k, form, df){
  
  # separate fold (test) data
  fold_dat <- df |> 
    filter(fold_random == k)
  
  # and training data
  train_dat <- df |> 
    filter(fold_random != k)
  
  # fit model with training data
  it_lm <- lm(form,
              data = train_dat)
  
  # generate predictions for the held-out fold data
  fold_preds <- augment(it_lm, newdata = fold_dat) |> 
    mutate(response = .fitted + .resid)

  # calculate R2 for the held-out fold data and save it
  r2_calc(obs = fold_dat$weight,
          pred = fold_preds)
}

k_fold_cv <- function(data, form, k=5) {
  n <- nrow(data)
  
  #split data into fold
  data <- data|> 
    mutate(fold_random = sample(rep_len(1:k, length.out = n),
                                size = n))
  
  #iteratively apply train/test to folds
  
  map_dbl(.x = 1:k,
          .f = ~ cv_it(k = .x, form = form, df = data))
  #average k across fold

  #k_fold_cv(data = ncbirths, form = weight ~ weeks, k = 5)
}
```