
```{r}
label: setup
library(tidyverse)
library(palmerpenguins)
#| message: false
```


```{r}
data(penguins)
```

# Exploring the Penguins Data

## Step 1: Barchart

7. Create a plot of the frequency of each penguin species in the data.

a. Use <https://excalidraw.com/> (or pen and paper, a tablet, etc.) to create a sketch for this plot. Label the aesthetics that will be needed.

b. Use `ggplot2` to create the plot you sketched above.

```{r}
#| label: barchart
ggplot(data = penguins,
       mapping = aes(x = species)) + 
  geom_bar() + 
  labs(x = "Species",
       y = "Frequency",
       title = "Frequency of Penguin Species")

```

<!-- Rotate roles! Project manager becomes coder, coder becomes computer, etc. -->

## Step 2: Histogram or Density Curve

8. Use `ggplot2` to plot the distribution of bill lengths for the penguins included in the dataset.

```{r}
#| label: histogram

ggplot(data = penguins,
       mapping = aes(x = bill_len)) + 
  geom_histogram()

```

## Step 3: Scatterplot

9. Use `ggplot2` to plot the relationship between the length of a penguin's bill (`bill_length_mm`) and the depth of their bill (`bill_depth_mm`).

```{r}
#| label: scatterplot
ggplot(data = penguins,
       mapping = aes(x = bill_len, 
                     y = bill_dep,
                     color = species)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```


## Step 4: Add a Trend Line

10. Add a **linear** trend line to the scatterplot you made above! 

<!-- Rotate roles! Project manager becomes coder, coder becomes computer, etc. -->

## Step 5: Adding A Categorical Variable

11. Building off of the plot you made in Step 4, add an aesthetic to differentiate the species of the penguins in the scatterplot by color for both the points *and* the trend line.

```{r}
#| label: scatterplot-with-one-categorical-variable


```

12. Edit your plot in (11) above so that the points are colored by species, but there is only one overall linear trend line.

```{r}
#| label: scatterplot-one-trend

#| label: scatterplot
ggplot(data = penguins,
       mapping = aes(x = bill_len, 
                     y = bill_dep,
                     )) + 
  geom_point(mapping = aes(color = species,
                           shape = island)) + 
  geom_smooth(method = "lm")
```


13. Building on your code from (11), add the location of the penguins (`island`) to your visualization. There is more than one method to address this, however, one method will more easiliy allow you to address the quesitons below.

```{r}
#| label: scatterplot-with-two-categorical-variables

count(penguins[,1])
```

## Canvas Quiz


Working as a team, use the plots you created to answer the following questions on Canvas.

1.  Which species of penguins is represented **least** in the Palmer Penguins data set?
Chinstrap
2. Which species had the weakest relationship between bill length and bill depth?
Adelie
3.  Which species of penguins are found on **every** island?
Adelie
4.  Which species of penguins are found **only** on Dream Island?
Chinstrap
5.  Which species of penguins are found **only** on Biscoe Island?
Gentoo


